---
title: "King-Rook vs. King-Pawn"
author: "Chinki"
date: "June 10, 2017"
output: word_document
---
King-Rook vs. King-Pawn
Source: https://archive.ics.uci.edu/ml/datasets/Chess+%28King-Rook+vs.+King-Pawn%29
A typical board-description is

f,f,f,f,f,f,f,f,f,f,f,f,l,f,n,f,f,t,f,f,f,f,f,f,f,t,f,f,f,f,f,f,f,t,t,n,won

The names of the features do not appear in the board-descriptions.
Instead, each feature correponds to a particular position in the
feature-value list.  For example, the head of this list is the value
for the feature "bkblk".  The following is the list of features, in
the order in which their values appear in the feature-value list:

[bkblk,bknwy,bkon8,bkona,bkspr,bkxbq,bkxcr,bkxwp,blxwp,bxqsq,cntxt,dsopp,dwipd,
 hdchk,katri,mulch,qxmsq,r2ar8,reskd,reskr,rimmx,rkxwp,rxmsq,simpl,skach,skewr,
 skrxp,spcop,stlmt,thrsk,wkcti,wkna8,wknck,wkovl,wkpos,wtoeg]

```{r}
#Reading the data
library(readxl)
Book1 <- read_excel("C:/Computational Statistics/Data Science workgroup/King/Book1.xlsx")

```

```{r}
#Getting structure
Book1$Column37=as.factor(Book1$Column37)
str(Book1)
```
```{r}
#Praportion of win and not win
prop.table(table(Book1$Column37))*100
```
chance of wining 52% and chance of not wining 48%.
```{r}
#Deleting missing values
Book1=na.omit(Book1)
#Training and test data
set.seed(134)
text=sample(3198,2500)
train=Book1[text,]
test=Book1[-text,]
train_label=train$Column37
test_label=test$Column37
```
```{r}
library(e1071)
nv=naiveBayes(Column37~.,data = train)
```
```{r}
pred=predict(nv,test)
pred
```
```{r}
library(gmodels)
CrossTable(pred,test_label)
```
```{r}
library(kernlab)
classification=ksvm(Column37~.,data=train,kernel="rbfdot")
```

